---
title: "Final Project Presentation"
subtitle: "Abby Sikora"
execute: 
  echo: false
  warning: false
format: 
  revealjs:
    embed-resources: true
---

```{r}
library(tidyverse)
library(readr)
library(here)
library(rvest)

#Iowa
url2 <- "https://www.espn.com/womens-college-basketball/team/stats/_/id/2294/iowa-hawkeyes"

h2 <- read_html(url2)

tab2 <- h2 |> html_nodes("table")

#stats with no names
iowa_df <- tab2[[4]] |> html_table(fill = TRUE)

#names table
iowa_df2 <- tab2[[3]] |> html_table(fill = TRUE)

iowa_stats <- bind_cols(iowa_df2, iowa_df)

#South Carolina

url3 <- "https://www.espn.com/womens-college-basketball/team/stats/_/id/2579/south-carolina-gamecocks"

h3 <- read_html(url3)

tab3 <- h3 |> html_nodes("table")

#stats with no names
sc_df <- tab3[[4]] |> html_table(fill = TRUE)

#names table
sc_df2 <- tab3[[3]] |> html_table(fill = TRUE)

sc_stats <- bind_cols(sc_df2, sc_df)

```


## Data Introduction

* 2 data sets, NCAA Women's basketball 2024
* Iowa - 14 players
* South Carolina - 12 players
* Both have 16 columns including stats like MIN(minutes), PTS(points), 3PM(3 pointers made), BLK(blocks) etc.

## Question?

I want to compare total season stats for both Iowa and South Carolina, see who their top players are, and just from some basic comparisons which team should've won the Championship


## Visualization 1

```{r}
sc_ppg <- sc_stats |> select(Name, PTS) |>
  arrange(desc(PTS))|>
  slice(2:12) |>
  mutate(Name = fct_reorder(Name, PTS))

sc_plot1 <- ggplot(data = sc_ppg, aes(x = Name,
                          y = PTS,
                          label = PTS)) +
  geom_point(color = "black") +
  geom_segment(aes(x = Name, xend = Name, y = 0, yend = PTS), color="red4") +
  coord_flip() +
  theme_minimal(base_size = 15) +
  theme(plot.background = element_rect(fill = "grey1"),
        axis.text = element_text(colour = "red3", size = rel(1)))

ggplotly(sc_plot1, tooltip = "label")
```

## Visualization 2

```{r}
library(plotly)

iowa_ppg <- iowa_stats |> select(Name, PTS) |>
  arrange(desc(PTS))|> 
  slice(2:13) |>
  mutate(Name = fct_reorder(Name, PTS))

iowa_plot1 <- ggplot(data = iowa_ppg, aes(x = Name,
                              y = PTS,
                              label = PTS)) +
  geom_point(color = "black") +
  geom_segment(aes(x = Name, xend = Name, y = 0, yend = PTS), color="yellow") +
  coord_flip() +
  theme_minimal(base_size = 15) +
  theme(plot.background = element_rect(fill = "grey1"),
        axis.text = element_text(colour = "yellow", size = rel(1))) 

ggplotly(iowa_plot1, tooltip = "label")
```
