---
title: "Final project"
format: html
---

```{r}
library(tidyverse)
library(rvest)
library(here)

url <- "https://www.ncaa.com/stats/basketball-women/d1/current/individual/102"

h <- read_html(url)

tab <- h |> html_nodes("table")
tab

wbb_df <- tab[[1]] |> html_table(fill = TRUE)

```

Variables in the dataset:

Rank: The rankings of highest points per game in descending order of the top 50 players
Name: Name of the player
Team: Name of the team the player is on
CL: Class Level (Fr., So., Jr., or Sr.)
Height: Height of the player
Position: (G - Guard, F - Forward, C - Center)
G: The number of games the player has played
FGM: Field Goals Made, the # of baskets each player has successfully scored on
3FG: 3 point field goals made by the athlete
FT: Free Throws made by the athlete
PTS: The number of total points scored by field goals per player
PPG: Average points per game scored by the player

Note: some of the players have "-" for rank, this means they have the same number of PPG as the person above them. 
The rank is only valid for PPG, all other stats are just extra included


Questions possible to explore: 
Maybe use another data set of viewership per year and see other connections for a part 2

visualizations: 
Histogram of ppg by player - maybe add plotly to see exact ppg

Distribution of class level

Something with total points

Boxplot of height

Does height correlate with ppg: scatter plot

```{r}
ggplot(data = wbb_df, aes(x = Height, y = PTS)) +
  geom_boxplot() +
  theme_minimal()
```











