---
title: "Final Edited"
author: "Abby Sikora"
format: 
  html:
    embed-resources: true
---
Question: 
Did South Carolina deserve the win over Iowa in the NCAA National Championship for women's basketball?

Iowa Dataset

```{r}
library(tidyverse)
library(readr)
library(here)
#Iowa
url2 <- "https://www.espn.com/womens-college-basketball/team/stats/_/id/2294/iowa-hawkeyes"

h2 <- read_html(url2)

tab2 <- h2 |> html_nodes("table")

#stats with no names
iowa_df <- tab2[[4]] |> html_table(fill = TRUE)

#names table
iowa_df2 <- tab2[[3]] |> html_table(fill = TRUE)

iowa_stats <- bind_cols(iowa_df2, iowa_df)

#South Carolina

url3 <- "https://www.espn.com/womens-college-basketball/team/stats/_/id/2579/south-carolina-gamecocks"

h3 <- read_html(url3)

tab3 <- h3 |> html_nodes("table")

#stats with no names
sc_df <- tab3[[4]] |> html_table(fill = TRUE)

#names table
sc_df2 <- tab3[[3]] |> html_table(fill = TRUE)

sc_stats <- bind_cols(sc_df2, sc_df)

```
THe data sets above are season statistics for women's NCAA basketball teams the Iowa Hawkeyes and t

These data sets includes variables:

Name: Name of the player
MIN: Minutes Per Game
FGM: Field Goals Made
FGA: Field Goals Attempted
FTM: Free Throws Made
FTA: Free Throws Attempted
3PM: 3-Pointers Made
3PA: 3-Pointers Attempted
PTS: Points
OR: Offensive Rebounds
DR: Defensive Rebounds
REB: Rebounds (Offensive and Defensive Total)
AST: Assists
TO: Turnovers
STL: Steals
BLK: Blocks


First, let's compare points between the teams. To make this data even, we will only look at the top ten players for amount of time playes (MIN) from each team because they have a different amount of players for this and the following few tables. 

```{r}
library(pander)

summary_iowa <- iowa_stats |> 
  arrange(desc(MIN)) |>
  slice(1:10) |>
  summarize(points_avg = mean(PTS))

summary_south_carolina <- sc_stats |>
  arrange(desc(MIN)) |>
  slice(1:10) |>
  summarize(points_avg = mean(PTS))

combined_summary <- bind_rows(
  mutate(summary_iowa, Team = "Iowa"),
  mutate(summary_south_carolina, Team = "South Carolina")) |>
  pander()
```
From this table, we see that Iowa has 27.4 more points on average for the season. To analyze this number further, I want to look at Field Goals Made, because the number of points doesn't tell us much on it's own.

Now let's look at Field Goals Made for each team.

```{r}
summary_iowa2 <- iowa_stats |> 
  arrange(desc(MIN)) |>
  slice(1:10) |>
  summarize(fg_made = mean(FGM))

summary_south_carolina2 <- sc_stats |>
  arrange(desc(MIN)) |>
  slice(1:10) |>
  summarize(fg_made = mean(FGM))

combined_summary2 <- bind_rows(
  mutate(summary_iowa2, Team = "Iowa"),
  mutate(summary_south_carolina2, Team = "South Carolina")) |>
  pander()
```

From this table, we can see that this time, South Carolina has a better number but not by much(<1). This tells us that although Iowa has a higher average of points total, the accuracy of the two teams is pretty similar when it comes to average field goals actually made. An explanation for the first table could be Iowa may have a higher number of 3 pointers, so I want to look at that next. 

```{r}
summary_iowa3 <- iowa_stats |> 
  arrange(desc(MIN)) |>
  slice(1:10) |>
  summarize(`3s_made` = mean(`3PM`))

summary_south_carolina3 <- sc_stats |>
  arrange(desc(MIN)) |>
  slice(1:10) |>
  summarize(`3s_made` = mean(`3PM`))

combined_summary3 <- bind_rows(
  mutate(summary_iowa3, Team = "Iowa"),
  mutate(summary_south_carolina3, Team = "South Carolina")) |>
  pander()
```
From this, we can see something that I had a feeling about from the previous tables. Iowa has almost double the amount of three pointers made than South Carolina. This tells us that the points average has a bigger margin of difference between the teams because Iowa simply scores higher value points more often than South Carolina.















